On the relation between metacognitive monitoring and domain knowledge

### 2019-11-19 ms.Rmd
wip

---
title             : "On the relation between metacognitive monitoring and domain knowledge"
shorttitle        : "Metacognitive resolution and domain knowledge"

author: 
  - name          : "Matti Vuorre"
    affiliation   : "1"
    corresponding : yes
    address       : "406 Schermerhorn Hall, 1190 Amsterdam Avenue, New York, NY 10027"
    email         : "vuorre@psych.columbia.edu"
  - name          : "Janet Metcalfe"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Columbia University"

authornote: |
  This research was supported, in part, by Institute of Education Science grant (R305A150467) awarded to JM. The authors are solely responsible for the content of this article. The data and analysis code supporting this manuscript are available at https://osf.io/np3cu/.

abstract: |
  Metacognitive resolution refers to people's ability to tell their correct answers apart from incorrect answers, and is theorized an important factor in learning. Resolution---the trial-by-trial association between response accuracy and metacognitive endorsement of that response's accuracy---is a distinct ability from knowledge in the domain of testing, and instead indicates a higher-order self-evaluation of one's cognitive states. It is therefore important that measures of resolution are independent of domain knowledge. We conducted six experiments that revealed a robust positive correlation between metacognitive resolution and performance in multiple-choice mathematics testing. A simulation study suggested that resolution metrics are increasingly negatively biased with decreasing performance, because multiple-choice tasks permit correct guessing. Finally, an experiment supported the guessing-related confound hypothesis: Resolution and performance were positively correlated in multiple-choice testing, but not in free-response testing. This study brings to light a previously underappreciated limitation in assessing metacognitive resolution and its relation to task performance when the criterion task involves guessing.

keywords          : "metacognition, confidence, measurement"
wordcount         : "`r wordcountaddin::word_count()`"

bibliography      : "references.bib"

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_word
---

```{r setup, include = FALSE}
source("functions.R")
library(papaja)
library(knitr)
library(Hmisc)
library(afex)
library(ggbeeswarm)
library(flextable)
library(patchwork)
library(broom)
library(scales)
library(tidyverse)
library(glue)
library(ordinal)
library(beepr)
library(parallel)
opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  cache = TRUE,
  dev = c("png", "pdf"),
  fig.ext = c("png", "pdf"),
  fig.retina = 2
)
options(mc.cores = parallel::detectCores())
theme_set(
  theme_linedraw(base_size = 8) + 
    theme(panel.grid = element_blank(), legend.position = "right")
)
# Function to format p values
pvalue <- function(x) {
  out <- scales::pvalue(x, add_p = T)
  out <- str_replace(out, "<", " < ")
  out <- str_replace(out, "=", " = ")
  out
}
```

@@

```{r math-example, out.width = "4cm", fig.cap = "Example mathematics problem from Experiment 1."}
include_graphics("example-problem-14.8.13.png")
```

@@

The students completed three pen-and-paper tests, separated by eight days of in-class instruction, as part of an after-school mathematics tutorial. Each test consisted of 20 four-alternative multiple-choice problems from the same pool of problems as in experiments 1A and B. Multiple problems were printed on each page of the test, but otherwise the test procedure was the same as in experiments 1A and B: Participants wrote their answers and confidence ratings---on the same scale as in the previous experiments---on a grid on the first page of the test. Participants completed the tests in groups of up to 25 students in a classroom supervised by a teacher, and had 45 minutes for each test.

@@

The participants used a website to view four-alternative multiple-choice mathematics problems. After logging on to the website, participants were presented one test problem at a time, such that the problem was shown on top of the screen, and the answer options below it. Participants clicked a small button next to their preferred response option, and then a "Submit" button to proceed to the next screen. There, they were prompted to rate their confidence in the accuracy of the preceding answer on a visual analog scale using the computer mouse. At the top of this screen the text "Sure correct or just guessing" was displayed, and a visual analog scale with endpoints labelled with "0%" on the left and "100%" was displayed below the text. In addition to the endpoints, the word "guessing" was displayed just below ¼ of the scale (without a tick mark), and "completely certain" just below the right endpoint. The word "CONFIDENCE" was printed below the scale and labels. Participants clicked on the scale to indicate their confidence, and then a "Submit" button to proceed to the next screen. After each confidence rating, participants could choose to receive feedback and/or watch instructional videos related to the problem, but those were unrelated to the current manuscript. Then, they moved on to the next problem.

In each tutorial session, the participants completed ten mathematics problems or made four errors, whichever occurred first. (They had the opportunity to redo problems which they had answered incorrectly, but those data are excluded from the current analyses.) The tests were completed on a website, and each participant completed the tests on an individual computer in a class of up to 25 students, under teacher supervision. 80 unique problems were included from the same pool as in the previous Experiments.

@@

Second, we used the signal detection theoretic measure *d*~a~ [@benjamin2008measurement; @MassonSourcesbiasGoodman2009], which measures the separation between two gaussian latent variables weighted by the ratio of their standard deviations. We calculated *d*~a~ with a linear regression of the zROC, as detailed by @benjamin2008measurement. However, that method ignores error on the x-axis [@HighamNoSpecialsignal2007; @KnoblauchModelingPsychophysicalData2012], and we therefore verified these results using more robust ordinal regression methods [@BurknerOrdinalRegressionModels2019]. The results were practically identical, so we focus on the former method's results because it is more common in the existing literature. In calculating the ROC based metrics, we added 1/K (K = number of confidence categories) to all rates to account for cells with zero observations [e.g. @HautusCorrectionsextremeproportions1995]. We also did the analyses without adjusting for empty cells. That led to many participants being dropped due to indeterminate measures, but otherwise the results were practically identical. For experiments 1D and 1F, where an analogue scale was used, an ordered scale with 6 categories was created.

@@

```{r math-dtab}
new_names <- tibble(
  key = c("zroc_auc_gamma", "gamma_old", "zroc_da", "phi", "zroc_auc", "trap_auc", "proportion_correct", "mean_rating"),
  Metric = c("Gamma (ROC)", "Gamma (pairs)", "da", "Phi", "AUC (zROC)", "AUC (trap)", "Proportion Correct", "Mean confidence")
)
tab <- math_metrics %>%
  gather(key, value, proportion_correct, zroc_auc_gamma, zroc_da, gamma_old) %>%
  group_by(key, Experiment) %>%
  group_modify(~ data.frame(t(smean.cl.boot(.x$value, B = 1000)))) %>%
  ungroup() %>%
  left_join(new_names) %>%
  right_join(math_subjects) %>%
  mutate_at(vars(Mean, Lower, Upper), ~ format(round(., 2), nsmall = 2)) %>%
  mutate(result = glue("{Mean} [{Lower}, {Upper}]")) %>%
  select(Experiment, N, Metric, result) %>%
  spread(Metric, result) %>%
  select(Experiment, N, `Proportion Correct`, everything())
kable(tab, caption = "Measures of mathematics performance and resolution in Experiment 1. Numbers indicate means with 95\\% CIs in square brackets.")
```

@@

```{r math-rfig, fig.cap = "Scatterplot of participants resolution metrics and mathematics proportions correct in Experiments 1A-F. Line of best fit and 95\\% CI are from an ordinary linear regression model."}
math_metrics %>%
  group_by(Experiment) %>% 
  gather(key, value, zroc_auc_gamma, zroc_da, gamma_old) %>%
  left_join(new_names) %>% 
  ggplot(aes(proportion_correct, value)) +
  geom_hline(yintercept = 0, lty = 2, size = .25) +
  scale_y_continuous(
    "Resolution",
    breaks = pretty_breaks()
  ) +
  scale_x_continuous(
    "Proportion correct",
    breaks = pretty_breaks()
  ) +
  geom_point(shape = 21, alpha = .5, size = .75) +
  geom_smooth(
    method = "lm", se = T,
    col = "black", size = .4
  ) +
  facet_grid(Metric ~ Experiment, scales = "free_y") +
  theme(aspect.ratio = 1)
```

@@

```{r math-rtab}
tab <- math_metrics %>%
  gather(key, value, zroc_da, zroc_auc_gamma, gamma_old) %>%
  left_join(new_names) %>%
  group_by(Experiment, Metric) %>% 
  group_modify(
    ~tidy(cor.test(.x$value, .x$proportion_correct, method = "kendall"))
  ) %>%
  mutate(p.value = pvalue(p.value)) %>%
  mutate(result = format(round(estimate, 2), nsmall = 2, trim = TRUE)) %>%
  mutate(result = glue("{result} ({p.value})")) %>%
  select(Experiment, Metric, result) %>%
  ungroup() %>%
  pivot_wider(names_from = c(Metric), values_from = result)
kable(
  tab, 
  caption = "Correlation between mathematics performance and three metrics of metacognitive resolution in Experiment 1."
)
```

@@

```{r sim-rfig, fig.cap = "Results of the simulation study at various levels of resolution (color), knowledge (*x*-axis), and guessing (columns), for *d*~a~ and gamma (rows)."}
sims_long %>% 
  # filter(d_gen %in% delta_bins[seq(1, length(delta_bins), 2)]) %>%
  rename(Guessing = p_guess) %>% 
  ggplot(aes(p_know, estimate, col=d_gen, group=d_gen)) +
  scale_color_viridis_c("Resolution") +
  geom_line(size=.6) +
  scale_x_continuous("Knowledge", breaks = pretty_breaks(5), limits = 0:1) +
  scale_y_continuous("Mean estimated resolution", breaks = pretty_breaks(5)) +
  facet_grid(Metric~Guessing, scales="free_y", labeller=label_both) +
  theme(legend.position = "bottom")
```

@@

Finally, we asked whether the participants' resolution was different from perfect if guessing was accounted for. We tested participants' *d*~a~s against the estimates from the simulation with Guessing = 0.25 and Resolution as the greatest value used in our simulations (2.5), at the level of the participants' performance, with a paired *t*-test. The simulated mean estimates of *d*~a~ from near-perfect resolution and Guessing = 0.25 were greater than the participants' observed *d*~a~s (`r res`): Participants did not have perfect metacognitive resolution. Further, better performing participants were further below near-perfect resolution than were poorer performing participants. Using gamma as the resolution metric yielded identical results.

### A note about calibration

Guesses may also impact measures of calibration, a different measure of metacognitive accuracy [@DunloskyMetacognition2009; @Lichtensteinthosewhoknow1977]. Because guesses artificially inflate measures of performance, a perfectly calibrated person's mean judgment---which aligns with their level of knowledge, not observed performance when some answers are guesses---will be lower than their inflated performance measure: Guessing will bias measures of calibration towards *under*confidence, such that the lowest-performing individuals will have the strongest guessing-induced underconfidence bias. One way to address this bias might be to estimate participants' knowledge using e.g. Abbott's formula (see above) instead of a simple proportion correct. As an extreme example, consider a participant in a 4AFC task who performs at chance (i.e. does not know the answer to any of the items) and accurately gives an average confidence judgment of 0%. Due to correct guessing, her observed percentage correct is 25%, leading to an observed underconfidence of 25%. However, a participant who truly knows the answer to all items, and knows that she knows (gives an average confidence rating of 100%), will not be impacted by correct guesses and thus will not show an underconfidence effect.

The guessing-induced underconfidence bias in calibration brings to mind a well-known mismatch between self-evaluations and skill: The Dunning-Kruger effect describes a phenomenon whereby, across a wide variety of tasks and contexts, individuals with poor task performance are prone to overestimating their skill [@KrugerUnskilledUnawareIt1999]. The reasons underlying the Dunning-Kruger effect are still debated, and some researchers have suggested that statistical artefacts and scale effects may partly explain the tendency of poorer performers to have inflated self-assessments [@FeldEstimatingrelationshipskill2017; @KrajcAreunskilledreally2008; @McIntoshWiseClarifyingrole2019]. However, when possible, guesses will lead to a bias in the opposite direction of the Dunning-Kruger effect. In studies on calibration and related phenomena, such as the Dunning-Kruger effect, researchers should be aware of the potentially biasing impact of lucky guesses on measures of calibration.

## Summary of simulation study

In summary, our simulation study suggested a general and important source of variation in resolution metrics in studies of metacognition using multiple-choice testing: The more opportunities there are for correct guesses---either due to number of response alternatives or to knowledge on the primary task---the more will measures of resolution be biased toward being too small. However, in order to examine this bias, we assumed a generative model of confidence ratings that relied on random guessing in the primary task. One concern with this model was that such random guessing, while intuitively appealing, may not be empirically plausible [e.g. @HighamNoSpecialsignal2007].

Therefore, we put the idea that resolution metrics are related to domain knowledge in multiple-choice tasks, possibly due to guessing, to an experimental test. In Experiment 2, we directly compared indices of metacognitive resolution, and their relation to domain skill, between free-response trials and multiple-choice trials.

@@

We recruited 100 participants through the Prolific (https://prolific.ac) online platform. `r tsex` participants were female, and age ranged from `r tage[1]` to `r tage[2]` years. We recruited participants currently residing in the US, and `r tfl` reported English as their first language. Participants were paid $3.64 for participating. All participants provided informed consent. The sample size was determined with a power simulation based on the results of the simulation study described above.

@@

We collected general knowledge questions from @TauberGeneralknowledgenorms2013 and from previous experiments conducted in our laboratory. Based on those previous studies, we selected 120 questions of varying degrees of difficulty, and found the most frequently endorsed commission errors to be used as the false response options in the 2AFC condition. We also screened the questions (and response alternatives) for potentially offensive or outdated material, and replaced unsuitable false response options with new ones. As an example, one question was "What was the name of the clown on the 'howdy doody' television show?". (Correct response: "Clarabell", incorrect alternative: "Bozo").

@@

As in Experiment 1, metacognitive resolution was quantified with *d*~a~, the ROC based gamma, and the concordant-discordant pairs gamma [@benjamin2008measurement; @HighamNewimprovedgamma2018; @Nelsoncomparisoncurrentmeasures1984]. These measures, proportions correct, and mean confidence ratings are summarised in Table \@ref(tab:trivia-dtab). First, this table shows that both proportion correct and mean confidence were significantly greater in the 2AFC condition. More importantly, all three resolution metrics were greater in the free response condition. This dramatic difference is surprising insofar as participants' knowledge across the two conditions was presumably equal. These results supported hypothesis i: Resolution measures were lower in the 2AFC task.

```{r trivia-dtab}
tmp1 <- trivia_metrics %>% 
  select(
    session, condition, proportion_correct, 
    mean_rating, gamma_old, zroc_da, zroc_auc_gamma
  ) %>% 
  gather(key, value, -session, -condition) %>% 
  spread(condition, value) %>% 
  mutate(difference = `2AFC`-Free) %>% 
  group_by(key) %>% 
  group_modify(~tidy(t.test(.$difference))) %>% 
  # mutate(p.value = pvalue(p.value), parameter = round(parameter, 0)) %>% 
  mutate_if(is.numeric, ~format(round(., 2), nsmall = 2)) %>%
  ungroup() %>% 
  transmute(
    key = key,
    result = glue("{estimate} [{conf.low}, {conf.high}]")
  ) %>% 
  left_join(new_names) %>% 
  select(-key) %>% 
  spread(Metric, result) %>% 
  add_column(condition = "2AFC - Free", .before = 1)
tmp2 <- trivia_metrics %>%
  # left_join(tmp_mc) %>% 
  gather(
    key, 
    value, 
    proportion_correct, mean_rating, gamma_old, zroc_auc_gamma, zroc_da,
  ) %>% 
  group_by(key, condition) %>%
  group_modify(~ data.frame(t(smean.cl.boot(.x$value, B = 1000)))) %>%
  ungroup() %>%
  left_join(new_names) %>%
  # mutate(Experiment = factor(exp)) %>%
  mutate_at(vars(Mean, Lower, Upper), ~ format(round(., 2), nsmall = 2)) %>%
  mutate(result = glue("{Mean} [{Lower}, {Upper}]")) %>%
  select(condition, Metric, result) %>%
  spread(Metric, result) %>%
  select(condition, `Proportion Correct`, `Mean confidence`, everything())
# ft <- flextable(tab)
# ft <- set_caption(tab, "Table 1. Descriptives of Study 1.")
# ft
bind_rows(tmp2, tmp1) %>% 
  kable(caption = "Measures of general information test performance and metacognition in Experiment 2. Numbers indicate means with 95\\% CIs in square brackets.")
```

We then turned to the second part of our hypothesis; that resolution would be spuriously correlated to performance in a 2AFC task. To answer, we correlated the three resolution metrics to performance, separately in the two conditions, again using the nonparametric tau correlation. The correlation was significant for each of the three resolution metrics in the 2AFC condition. However, it was much smaller in the free response condition, where it was only significant for the concordant-discordant pairs based gamma. These scatterplots are shown in Figure \@ref(fig:trivia-rfig).

```{r trivia-ctab}
tmp <- trivia_metrics %>%
  gather(key, value, gamma_old, zroc_da, zroc_auc_gamma) %>%
  left_join(new_names) %>%
  group_by(condition, Metric)
tab <- group_modify(
  tmp,
  ~tidy(cor.test(.x$value, .x$proportion_correct, method = "kendall"))
) %>%
  mutate(p.value = pvalue(p.value)) %>%
  mutate(result = format(round(estimate, 2), nsmall = 2, trim = TRUE)) %>%
  mutate(result = glue("{result} ({p.value})")) %>%
  select(condition, Metric, result) %>%
  ungroup() %>%
  pivot_wider(names_from = c(Metric), values_from = result)
kable(
  tab, 
  caption = "Correlation between general information test performance and three metrics of metacognitive resolution in Experiment 2."
  )
```

>>>>

---
title: "On the relation between metacognitive monitoring and domain knowledge"
shorttitle: "Metacognitive resolution and domain knowledge"

author: 
  - name: "Matti Vuorre"
    affiliation: "1"
    corresponding: yes
    address: "406 Schermerhorn Hall, 1190 Amsterdam Avenue, New York, NY 10027"
    email: "vuorre@psych.columbia.edu"
  - name: "Janet Metcalfe"
    affiliation: "1"

affiliation:
  - id: "1"
    institution: "Columbia University"

authornote: |
  This research was supported, in part, by Institute of Education Science grant (R305A150467) awarded to JM. The authors are solely responsible for the content of this article. The data and code supporting this manuscript are available at https://osf.io/np3cu/.

abstract: |
  Metacognitive resolution refers to people's ability to tell their correct answers apart from incorrect answers, and is theorized an important factor in learning. Resolution---the trial-by-trial association between response accuracy and metacognitive endorsement of that response's accuracy---is a distinct ability from knowledge in the domain of testing, and instead indicates a higher-order self-evaluation. It is therefore important that measures of resolution are independent of domain knowledge. Six multiple-choice mathematics experiments revealed a robust positive correlation between resolution and performance. A simulation study suggested that correct guesses afforded by multiple-choice testing necessarily lead to a correlation between resolution and performance. Finally, an experiment supported this hypothesis: Resolution and performance were positively correlated in multiple-choice, but not in free-response testing. These results suggest that the "unskilled and unaware" effect, whereby poorly performing individuals appear unaware of their lack of ability, may be artifactual in multiple-choice tasks that define "awareness" as metacognitive resolution.

keywords: "metacognition, confidence, measurement, unskilled and unaware"
wordcount: "`r wordcountaddin::word_count()`"

bibliography: "references.bib"

floatsintext: yes
figurelist: no
tablelist: no
footnotelist: no
linenumbers: no
mask: no
draft: no

documentclass: "apa6"
classoption: "man"
output: 
  papaja::apa6_pdf:
    keep_tex: false
  papaja::apa6_word:
    default 
---

```{r setup, include = FALSE}
source("functions.R")
library(papaja)
library(knitr)
library(Hmisc)
library(afex)
library(ggbeeswarm)
library(kableExtra)
library(patchwork)
library(broom)
library(scales)
library(tidyverse)
library(glue)
library(ordinal)
library(beepr)
library(parallel)
opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  cache = TRUE, 
  fig.align = 'center'
)
options(mc.cores = parallel::detectCores())
theme_set(
  theme_linedraw(base_size = 10) + 
    theme(panel.grid = element_blank(), legend.position = "right")
)
# Function to format p values
pvalue <- function(x) {
  out <- scales::pvalue(x, add_p = T)
  out <- str_replace(out, "<", " < ")
  out <- str_replace(out, "=", " = ")
  out
}
```

@@

```{r math-example, fig.cap = "Example mathematics problem from Experiment 1."}
include_graphics("example-problem-14.8.13.png")
```

@@

The students completed three pen-and-paper tests, separated by eight days of in-class instruction, as part of an after-school mathematics tutorial. Each test consisted of 20 four-alternative multiple-choice problems from the same pool of problems as in Experiments 1A and B. Multiple problems were printed on each page of the test, but otherwise the test procedure was the same as in Experiments 1A and B: Participants wrote their answers and confidence ratings---on the same scale as in the previous experiments---on a grid on the first page of the test. Participants completed the tests in groups of up to 25 students in a classroom supervised by a teacher, and had 45 minutes for each test.

@@

The participants used a website to view four-alternative multiple-choice mathematics problems. After logging on to the website, participants were presented one test problem at a time, such that the problem was shown on top of the screen, and the answer options below it. Participants clicked a small button next to their preferred response option, and then a "Submit" button to proceed to the next screen. There, they were prompted to rate their confidence in the accuracy of the preceding answer on a visual analog scale using the computer mouse. At the top of this screen the text "Sure correct or just guessing" was displayed. Below the text, a visual analog scale was displayed with endpoints labelled with "0%" on the left and "100%" on the right. In addition to the endpoints, the word "guessing" was displayed just below ¼ of the scale (without a tick mark), and "completely certain" just below the right endpoint. The word "CONFIDENCE" was printed below the scale and labels. Participants clicked on the scale to indicate their confidence, and then a "Submit" button to proceed to the next screen. After each confidence rating, participants could choose to receive feedback and/or watch instructional videos related to the problem, but those were unrelated to the current manuscript. Then, they moved on to the next problem.

In each tutorial session, the participants completed ten mathematics problems or made four errors, whichever occurred first. (They had the opportunity to redo problems which they had answered incorrectly, but those data are excluded from the current analyses.) The tests were completed on a website, and each participant completed the tests on an individual computer in a class of up to 25 students, under teacher supervision. Eighty unique problems were included from the same pool as in the previous Experiments.

@@

Second, we used the signal detection theoretic measure *d*~a~ [@benjamin2008measurement; @MassonSourcesbiasGoodman2009], which measures the separation between two gaussian latent variables weighted by the ratio of their standard deviations. We calculated *d*~a~ with a linear regression of the zROC, as detailed by @benjamin2008measurement. However, that method ignores error on the x-axis [@HighamNoSpecialsignal2007; @KnoblauchModelingPsychophysicalData2012], and we therefore verified these results using more robust ordinal regression methods [@BurknerOrdinalRegressionModels2019]. The results were practically identical, so we focus on the former method's results because it is more common in the existing literature. In calculating the ROC based metrics, we added 1/K (K = number of confidence categories) to all rates to account for cells with zero observations [e.g. @HautusCorrectionsextremeproportions1995]. We also did the analyses without adjusting for empty cells. That led to many participants being dropped due to indeterminate measures, but otherwise the results were practically identical. For Experiments 1D and 1F, where an analogue scale was used, an ordered scale with 6 categories was created.

@@

```{r math-dtab}
new_names <- tibble(
  key = c("zroc_auc_gamma", "gamma_old", "zroc_da", "phi", "zroc_auc", "trap_auc", "proportion_correct", "mean_rating"),
  Metric = c("Gamma (ROC)", "Gamma (pairs)", "da", "Phi", "AUC (zROC)", "AUC (trap)", "Proportion Correct", "Mean confidence")
)
tab <- math_metrics %>%
  gather(key, value, proportion_correct, zroc_auc_gamma, zroc_da, gamma_old) %>%
  group_by(key, Experiment) %>%
  group_modify(~ data.frame(t(smean.cl.boot(.x$value, B = 1000)))) %>%
  ungroup() %>%
  left_join(new_names) %>%
  right_join(math_subjects) %>%
  mutate_at(vars(Mean, Lower, Upper), ~ format(round(., 2), nsmall = 2)) %>%
  mutate(result = glue("{Mean} [{Lower}, {Upper}]")) %>%
  select(Experiment, N, Metric, result) %>%
  spread(Metric, result) %>%
  select(Experiment, N, `Proportion Correct`, everything())
kable(
  tab, 
  booktabs = TRUE,
  caption = "Measures of mathematics performance and resolution in Experiment 1. Numbers indicate means with 95\\% CIs in square brackets."
  )
```

@@

```{r math-rfig, fig.cap = "Scatterplot of participants resolution metrics and mathematics proportions correct in Experiments 1A-F. Line of best fit and 95\\%CI are from an ordinary linear regression model."}
math_metrics %>%
  group_by(Experiment) %>% 
  gather(key, value, zroc_auc_gamma, zroc_da, gamma_old) %>%
  left_join(new_names) %>% 
  ggplot(aes(proportion_correct, value)) +
  geom_hline(yintercept = 0, lty = 2, size = .25) +
  scale_y_continuous(
    "Resolution",
    breaks = pretty_breaks()
  ) +
  scale_x_continuous(
    "Proportion correct",
    breaks = pretty_breaks()
  ) +
  geom_point(shape = 21, alpha = .5, size = .75) +
  geom_smooth(
    method = "lm", se = T,
    col = "black", size = .4
  ) +
  facet_grid(Metric ~ Experiment, scales = "free_y") +
  theme(aspect.ratio = 1)
```

@@

```{r math-rtab}
tab <- math_metrics %>%
  gather(key, value, zroc_da, zroc_auc_gamma, gamma_old) %>%
  left_join(new_names) %>%
  group_by(Experiment, Metric) %>% 
  group_modify(
    ~tidy(cor.test(.x$value, .x$proportion_correct, method = "kendall"))
  ) %>%
  mutate(p.value = pvalue(p.value)) %>%
  mutate(result = format(round(estimate, 2), nsmall = 2, trim = TRUE)) %>%
  mutate(result = glue("{result} ({p.value})")) %>%
  select(Experiment, Metric, result) %>%
  ungroup() %>%
  pivot_wider(names_from = c(Metric), values_from = result)
kable(
  tab, 
  booktabs = TRUE,
  caption = "Correlation between mathematics performance and three metrics of metacognitive resolution in Experiment 1."
)
```

@@

```{r sim-rfig, fig.cap = "Results of the simulation study at various levels of resolution (color), knowledge, and guessing, for da and gamma."}
sims_long %>% 
  # filter(d_gen %in% delta_bins[seq(1, length(delta_bins), 2)]) %>%
  rename(Guessing = p_guess) %>% 
  ggplot(aes(p_know, estimate, col=d_gen, group=d_gen)) +
  scale_color_viridis_c("Resolution") +
  geom_line(size=.6) +
  scale_x_continuous("Knowledge", breaks = pretty_breaks(5), limits = 0:1) +
  scale_y_continuous("Mean estimated resolution", breaks = pretty_breaks(5)) +
  facet_grid(Metric~Guessing, scales="free_y", labeller=label_both) +
  theme(legend.position = "bottom")
```

@@

We then asked whether resolution in Experiments 1A-F was different from perfect if guessing was accounted for. We tested participants' *d*~a~s against the estimates from the simulation with Guessing = 0.25 and Resolution as the greatest value used in our simulations (2.5), at the level of the participants' performance, with a paired *t*-test. The simulated mean estimates of *d*~a~ from near-perfect resolution and Guessing = 0.25 were greater than the participants' observed *d*~a~s (`r res`): Participants did not have perfect metacognitive resolution. Further, better performing participants were further below near-perfect resolution than were poorer performing participants. Using gamma as the resolution metric yielded identical results.

In summary, our simulation study suggested that in studies of metacognition using multiple-choice testing, metrics of resolution will necessarily correlate with primary task performance. However, we assumed a generative model of confidence that relied on random guessing in the primary task. One concern with this model is that such random guessing, while intuitively appealing, may not be empirically plausible [e.g. @HighamNoSpecialsignal2007]. In fact, one study using SAT (five alternative multiple-choice) materials found that a signal detection theoretic (SDT) model described performance better than did a high-threshold model (i.e. model with random guessing; @HighamNoSpecialsignal2007). Specifically, high-threshold theory (as implied in our simulations) predicts that accuracy at the lowest level of confidence should not drop below 1/K (K = number of response alternatives). In contrast, as found in that study and predicted by SDT, accuracy was below 20% for low confidence responses, thus discrediting the random guessing assumption [@HighamNoSpecialsignal2007].

```{r accuracy-at-confidence, eval = TRUE}
tmp <- bind_rows(
  math %>% 
    filter(Experiment %in% c("1A", "1B", "1C")) %>% 
    group_by(Experiment, subject) %>% 
    mutate(con = bin_ratings(con, 1, 6, 6)) %>% 
    group_by(Experiment, subject, con) %>% 
    summarise(p = mean(accuracy)),
  math %>% 
    filter(Experiment %in% c("1D", "1F")) %>% 
    group_by(Experiment, subject) %>% 
    mutate(con = bin_ratings(con, 0, 1, 6)) %>% 
    group_by(Experiment, subject, con) %>% 
    summarise(p = mean(accuracy)),
  math %>% 
    filter(Experiment %in% c("1E")) %>% 
    group_by(Experiment, subject) %>% 
    mutate(con = bin_ratings(con, 0, 10, 6)) %>% 
    group_by(Experiment, subject, con) %>% 
    summarise(p = mean(accuracy))
) %>% 
  left_join(select(math_metrics, Experiment, subject, proportion_correct)) %>% 
  ungroup() %>% 
  mutate(Performance = cut2(proportion_correct, g = 3))
res <- tmp %>% 
  filter(con==1) %>% 
  group_by(Performance) %>% 
  group_modify(~tidy(t.test(.$p, mu = 0.25))) %>% 
  glue_data("{round(estimate*100, 1)}%, 95%CI = [{round(conf.low*100, 1)}, {round(conf.high*100, 1)}]")

# tmp %>%   
#   ggplot(aes(con, p, col = Performance)) +
#   scale_y_continuous(
#     "Proportion correct",
#     limits = 0:1
#   ) +
#   labs(x = "Confidence") +
#   stat_summary(geom = "line", aes(group=Performance)) +
#   stat_summary()
```

To begin to address this issue, we took the data of Experiments 1A-F, and asked if performance did indeed drop below 25% at the lowest level of confidence. We did this analysis separately for performance tertiles of participants, because good performing students should necessarily guess less than poor performers. In support of the guessing hypothesis, none of the performance tertiles had performance below 25% at the lowest level of confidence (Low performers: `r res[1]`, medium performers: `r res[2]`, high performers: `r res[3]`). In fact, as can be seen, the lowest performers were almost exactly at 25% at the lowest level of confidence. Nevertheless, these analyses were preliminary, and a stronger test of our hypothesis would be to experimentally contrast tasks where guessing was possible and where it was not. Therefore, we put the idea that resolution metrics are related to domain knowledge in multiple-choice tasks, possibly due to guessing, to an experimental test. In Experiment 2, we directly compared indices of metacognitive resolution, and their relation to domain skill, between free-response trials and multiple-choice trials.

@@

We recruited 100 participants through the Prolific (https://prolific.co) online platform. `r tsex` participants were female, and age ranged from `r tage[1]` to `r tage[2]` years. We recruited participants currently residing in the US, and `r tfl` reported English as their first language. All participants provided informed consent, and were paid $3.64. We determined the sample size with a power simulation based on the results of the simulation study described above, such that we would achieve 90% power for detecting a difference in the resolution-performance correlations between 2AFC and free response conditions.

@@

We collected general knowledge questions from @TauberGeneralknowledgenorms2013 and from previous experiments conducted in our laboratory. Based on those previous studies, we selected 120 questions of varying degrees of difficulty, and found the most frequently endorsed commission errors to be used as the false response options in the 2AFC condition. We also screened the questions (and response alternatives) for potentially offensive or outdated material, and replaced unsuitable false response options with new ones. As an example, one question was "What college or university, founded in 1636, is the oldest in the united states?". (Correct response: "Harvard", incorrect alternative: "Columbia").

@@

As in Experiment 1, metacognitive resolution was quantified with *d*~a~, the ROC based gamma, and the concordant-discordant pairs gamma [@benjamin2008measurement; @HighamNewimprovedgamma2018; @Nelsoncomparisoncurrentmeasures1984]. These measures, proportions correct, and mean confidence ratings are summarised in Table \@ref(tab:trivia-dtab). First, proportion correct and mean confidence were significantly greater in the 2AFC condition. More importantly, all three resolution metrics were greater in the free response condition. This dramatic difference is surprising insofar as participants' knowledge across the two conditions was presumably equal. These results supported hypothesis i: Resolution measures were lower in the 2AFC task.

```{r trivia-dtab}
tmp1 <- trivia_metrics %>% 
  select(
    session, condition, proportion_correct, 
    mean_rating, gamma_old, zroc_da, zroc_auc_gamma
  ) %>% 
  gather(key, value, -session, -condition) %>% 
  spread(condition, value) %>% 
  mutate(difference = `2AFC`-Free) %>% 
  group_by(key) %>% 
  group_modify(~tidy(t.test(.$difference))) %>% 
  # mutate(p.value = pvalue(p.value), parameter = round(parameter, 0)) %>% 
  mutate_if(is.numeric, ~format(round(., 2), nsmall = 2)) %>%
  ungroup() %>% 
  transmute(
    key = key,
    result = glue("{estimate} [{conf.low}, {conf.high}]")
  ) %>% 
  left_join(new_names) %>% 
  select(-key) %>% 
  spread(Metric, result) %>% 
  add_column(condition = "2AFC - Free", .before = 1)
tmp2 <- trivia_metrics %>%
  gather(
    key, 
    value, 
    proportion_correct, mean_rating, gamma_old, zroc_auc_gamma, zroc_da,
  ) %>% 
  group_by(key, condition) %>%
  group_modify(~ data.frame(t(smean.cl.boot(.x$value, B = 1000)))) %>%
  ungroup() %>%
  left_join(new_names) %>%
  mutate_at(vars(Mean, Lower, Upper), ~ format(round(., 2), nsmall = 2)) %>%
  mutate(result = glue("{Mean} [{Lower}, {Upper}]")) %>%
  select(condition, Metric, result) %>%
  spread(Metric, result) %>%
  select(condition, `Proportion Correct`, `Mean confidence`, everything())
bind_rows(tmp2, tmp1) %>% 
  kable(
    booktabs = TRUE,
    caption = "Measures of general information test performance and metacognition in Experiment 2. Numbers indicate means with 95\\% CIs in square brackets."
  ) %>% 
  kable_styling(latex_options = "scale_down")
```

We then turned to the second part of our hypothesis; that resolution would be spuriously correlated to performance in a 2AFC task. We correlated the three resolution metrics to performance in each conditions, using the nonparametric tau correlation. The correlation was significant for all three resolution metrics in the 2AFC condition (Table \@ref(tab:trivia-ctab)). However, it was much smaller in the free response condition, where it was only significant for the concordant-discordant pairs based gamma (Figure \@ref(fig:trivia-rfig)).

```{r trivia-rfig, fig.cap = "Scatterplot of participants' resolution metrics and proportions correct across the two conditions in Experiment 2. Lines of best fit and 95\\% CIs are from an ordinary linear regression model."}
trivia_metrics %>%
  gather(key, value, zroc_auc_gamma, zroc_da, gamma_old) %>%
  left_join(new_names) %>% 
  ggplot(aes(proportion_correct, value)) +
  geom_hline(yintercept = 0, lty = 2, size = .25) +
  scale_y_continuous(
    "Resolution",
    breaks = pretty_breaks()
  ) +
  scale_x_continuous(
    "Proportion correct",
    breaks = pretty_breaks()
  ) +
  geom_point(shape = 21, alpha = .5, size = 1) +
  geom_smooth(method = "lm", col = "black", size = .7) +
  facet_grid(Metric ~ condition, scales = "free_y") +
  theme(aspect.ratio = .7)
```
